---
title: "Analyses of species distributions in peatlands"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
  word_document:
    toc: yes
    toc_depth: '4'
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(tibble.width = Inf)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r Define ggplot themes and palettes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(panel.background = element_rect(fill = "white")) +
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+ 
    theme(panel.background = element_rect(fill = "white"))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
```

# Load packages

```{r load packages}
library(tidyverse)
library(readxl)
library(knitr)
library(ggeffects)
library(car)
library(glmmTMB)
library(ggthemes)
library(gridExtra)
library(performance)
library(ggplot2)
library(vegan)
library(rdacca.hp)
library(egg)
library(cowplot)
library(BiodiversityR)
library(ggrepel)
library(sjPlot)
library(DHARMa)
library(extrafont)
library(caret)
library(ggord)
library(patchwork)
```

# Data preparation

## Read data from Excel file

This reads the data from the sheet "SDM Data" in the Excel file "Modelling_SDM_species_data_AV.xlsx". Note that you need to change the path to the folder where you have the Excel file

```{r}
data_peat<-read_excel("Modelling_SDM_species_data_DLM_par.xlsx",
                      sheet="DLMC")

```

## Have a look at the data

```{r}
data_peat
```


## Convert some variables to factors

Convert some variables (those that are Y/N or 0/1) to factors.

```{r}
data_peat<-data_peat%>%
  mutate(fen=as.factor(fen),imp_temp=as.factor(imp_temp),
         nutrient=as.factor(nutrient),fire=as.factor(fire),dry=as.factor(dry)) 
```

## Convert moist to numeric

```{r}
data_peat<-data_peat%>%
  mutate(moist=as.numeric(moist))
```

## Create presence variables

Create new columns with presence/absence data. For each column, we create a column that starts with "pres_". This column will be 0 if abundance of the species is 0, and 1 otherwise

```{r}
data_peat<-data_peat%>%
  mutate(pres_tot_Sphagnum=ifelse(tot_Sphagnum>0,1,0),
         pres_Erio=ifelse(Erio>0,1,0), 
         pres_Carex=ifelse(Carex>0,1,0),
         pres_Erica=ifelse(Erica>0,1,0)) 
```

```{r}
data_peat
```

# Look at the distribution of some variables

Histogram for total Sphagnum abundance

```{r}
hist(data_peat$tot_Sphagnum)
hist(data_peat$pres_Erica)
hist(data_peat$pres_tot_Sphagnum)
hist(data_peat$pres_Carex)
```

Histogram for total Sphagnum presence (you can see that this is either 0 = absence or 1 = presence)

```{r}
hist(data_peat$pres_tot_Sphagnum)
```

Histogram for total Erio presence

```{r}
hist(data_peat$pres_Erio)
```

There seems to be very few cases where pres_Erio=0. We can see them.

```{r}
data_peat%>%filter(pres_Erio==0)
```

# Check the correlations among your independent variables

Construct a correlation matrix with all your (numeric) independent variables

```{r}
cor(data_peat%>%
      select(age,temp,moist, par),    # Select the 3 numeric independent variables
    use="pairwise.complete.obs") 
```

# Sphagnum species composition

## Distance-based redundancy analysis  

Data for ordination: 

```{r}
data_ordination<-data_peat %>%
  filter_at(vars(Medium:Tenellum), 
            all_vars(!is.na(.)))%>% # Remove rows with all NAs
  filter_at(vars(Medium:Tenellum),
            any_vars(.>0))%>% # Remove rows with all zeros
  filter(!is.na(age)&!is.na(temp)&!is.na(moist)&!is.na(nutrient)&!is.na(fire)&
           !is.na(dry))%>% # Remove rows with NA in predictors
  rename(Cuspidata=`Cuspidatum`) #Rename to avoid problems
```

Distance-based redundancy analysis (db-RDA) with Bray-Curtis distance.

Calculate ordination:

```{r}
ordination<-capscale(data_ordination[10:17]~temp+moist+nutrient+fire+dry, 
                     data = data_ordination, distance="bray") 
```

Result of the ordination:

```{r}
ordination
```

Proportion explained by each ordination axis. CAP1-CAP6 are the "constrained" axes, explained by the environmental variables. MDS1-MDS34 are the "unconstrained" axes.

```{r}
eigenvals(ordination) %>% summary()
```

Test significance of the ordination with Monte Carlo permutation test.

For the whole model:

```{r}
anova (ordination, permutations = 999) 
```

For each explanatory variable (with all the others used as covariables, independently from their order in the model):

```{r}
anova (ordination, by = 'margin', permutations = 999)
```

For each axis:

```{r}
anova (ordination, by = 'axis', permutations = 999)
```

## Hierarchical partitioning

```{r}
dist_matrix<-vegdist(data_ordination[10:17],method="bray")
```

```{r}
hierpart<-rdacca.hp(dist_matrix,data_ordination[c(19,21:24)],method = "dbRDA",
          type ="adjR2",scale = F,add = T,sqrt.dist = T)
hierpart
```

## VIF of the ordination

```{r}
vif.cca(ordination)
```

## Figure 3: Ordination plot + hierarchical partitioning

Ordination plot with vegan:

```{r}
ordination_plot<-ordiplot(ordination,display = c('species', 'sites', 'bp'))
```

Extract information on the locations of sites (circles in the ordiplot) via function sites.long. Information on characteristics of the sites is added via the argument of env.data.

Extract information on species with function species.long.

```{r}
sites.long <- sites.long(ordination_plot, env.data=data_ordination)
species.long <- species.long(ordination_plot)
head(sites.long)
species.long
```

```{r}
vectors.long  <- data.frame(ordination_plot$biplot[,1:2])
row.names(vectors.long)<-c("Temperature","Locally dry","Nutrient input = 1",
                            "Fire = 1","Regionally dry = 1")
```

```{r}
ordi_plot1 <- ggplot() + 
  geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +
  xlab("Axis 1 (67.6%)") + ylab("Axis 2 (23.4%)") +  
  scale_x_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
  scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +  
  geom_point(data=sites.long,aes(x=axis1, y=axis2),
             size=3,alpha=0.3,shape=16,color="darkturquoise") + 
  geom_point(data=species.long,aes(x=axis1, y=axis2),
             size=2,shape=1,color="brown2") +
  geom_text_repel(data=species.long,aes(x=axis1, y=axis2, label=labels),
                  colour="brown2",size=3,family="serif")+
  geom_segment(data=vectors.long, aes(x=0, y=0, xend=CAP1*2, yend=CAP2*2), 
               color="black", size=0.5,arrow=arrow(length=unit(0.02,"npc"))) +
  geom_text_repel(data=vectors.long[1,],
                  aes(x=CAP1*2.15,y=CAP2*5,
                      label=rownames(vectors.long[1,])), 
                  cex=3,direction="both",segment.size=0.25,color="black",
                  family="serif") +
  geom_text_repel(data=vectors.long[2,],
                  aes(x=CAP1*2.6,y=CAP2*2.3,
                      label=rownames(vectors.long[2,])), 
                  cex=3,direction="both",segment.size=0.25,color="black",
                  family="serif") +
  geom_text_repel(data=vectors.long[3,],
                  aes(x=CAP1*2.5,y=CAP2*2.5,
                      label=rownames(vectors.long[3,])), 
                  cex=3,direction="both",segment.size=0.25,color="black",
                  family="serif") +
  geom_text_repel(data=vectors.long[4,],
                  aes(x=CAP1*2.4,y=CAP2*2.1,
                      label=rownames(vectors.long[4,])), 
                  cex=3,direction="both",segment.size=0.25,color="black",
                  family="serif") +
  geom_text_repel(data=vectors.long[5,],
                  aes(x=CAP1*(-0.5),y=CAP2*2.4,
                      label=rownames(vectors.long[5,])), 
                  cex=3,direction="both",segment.size=0.25,color="black",
                  family="serif") +
  geom_text_repel(data=vectors.long[6,],
                  aes(x=CAP1*3,y=CAP2*2,
                      label=rownames(vectors.long[6,])), 
                  cex=3,direction="both",segment.size=0.25,color="black",
                  family="serif") +
  my_theme()
ordi_plot2<-ggplot(data.frame(hierpart$Hier.part)%>%
         rownames_to_column(var="variable")%>%
         mutate(variable=c("Temperature","Locally dry","Nutrient input","Fire",
                           "Regionally dry"))%>%
         mutate(variable=factor(variable,
                                levels=c("Temperature","Locally dry",
                                         "Nutrient input","Fire",
                                         "Regionally dry"))),
       aes(x=variable,y=I.perc...))+geom_bar(stat = "identity")+
  my_theme()+xlab(NULL)+ylab("Individual effect to R square (%)")
ordi_plot<-cowplot::plot_grid(ordi_plot1,ordi_plot2,ncol=2,labels=c("A)","B)"),
                   label_fontfamily="serif",align="hv")
ordi_plot  
ggsave(filename="figures/fig3.tiff",plot=ordi_plot,
       width=30,height=12,units="cm",dpi=300)
```


# Paleo-species distribution models

## Convert abundances to proportions between 0 and 1

```{r}
data_peat<-data_peat%>%
  mutate(tot_Sphagnum_prop=tot_Sphagnum/100)

data_peat<-data_peat%>%
  mutate(Erica_prop=Erica/100)

data_peat<-data_peat%>%
  mutate(Erio_prop=Erio/100)

data_peat<-data_peat%>%
  mutate(Carex_prop=Carex/100)

data_peat<-data_peat%>%
  mutate(par_prop=par/100)
```

## Models without fen and wihout interactions

### Species groups

#### Total Sphagnum 

```{r}
mod_abund_tot_Sphagnum<-glmmTMB(tot_Sphagnum_prop~ 
                                  temp+moist+
                                  nutrient+fire+dry, 
                                family="beta_family", 
                                ziformula=~.,
                                data=data_peat)    
summary(mod_abund_tot_Sphagnum)
check_collinearity(mod_abund_tot_Sphagnum)
```

#### Eriophorum

```{r}
mod_abund_Erio_nozi<-glmmTMB(Erio_prop~  
                               temp+moist+
                               nutrient+fire+dry,  
                             family="beta_family", 
                             ziformula=~0,       
                             data=subset(data_peat,Erio_prop>0)) 
summary(mod_abund_Erio_nozi)
check_collinearity(mod_abund_Erio_nozi)
```

#### Ericaceae

```{r}
mod_abund_Erica<-glmmTMB(Erica_prop~ 
                          temp+moist+
                          nutrient+fire+dry, 
                        family="beta_family", 
                        ziformula=~.,
                        data=data_peat)  
summary(mod_abund_Erica)
check_collinearity(mod_abund_Erica)
```

#### Carex

```{r}
data_peat<-data_peat%>%
  mutate(Carex_prop=Carex/100)
mod_abund_Carex<-glmmTMB(Carex_prop~  
                          temp+moist+
                          nutrient+fire+dry, 
                        family="beta_family",
                        ziformula=~.,
                        data=data_peat) 
summary(mod_abund_Carex)
check_collinearity(mod_abund_Carex)
```

### Individual Sphagnum species

#### Medium 

```{r}
mod_abund_Medium<-glmmTMB(Medium~  
                               temp+moist+
                               nutrient+fire+dry, 
                             family="beta_family",
                             ziformula=~.,         
                             data=data_peat) 
summary(mod_abund_Medium)
check_collinearity(mod_abund_Medium)
```

#### Fuscum 

```{r}
mod_abund_Fuscum<-glmmTMB(Fuscum~    
                               temp+moist+
                               nutrient+fire+dry,
                             family="beta_family", 
                             ziformula=~.,         
                             data=data_peat) 
summary(mod_abund_Fuscum)
check_collinearity(mod_abund_Fuscum)
```

#### Rubellum

```{r}
mod_abund_Rubellum<-glmmTMB(Rubellum~   
                               temp+moist+
                               nutrient+fire+dry, 
                             family="beta_family",
                             ziformula=~.,         
                             data=data_peat)
summary(mod_abund_Rubellum)
check_collinearity(mod_abund_Rubellum)
```

#### Cuspidata

```{r}
mod_abund_Cuspidatum<-glmmTMB(Cuspidatum~    
                               temp+moist+
                               nutrient+fire+dry,  
                             family="beta_family", 
                             ziformula=~.,         
                             data=data_peat) 
summary(mod_abund_Cuspidatum)
check_collinearity(mod_abund_Cuspidatum)
```

### Tests for temporal autocorrelation

#### Species groups

```{r}
testTemporalAutocorrelation(simulateResiduals(mod_abund_tot_Sphagnum),
                            subset(data_peat,
                                   !is.na(tot_Sphagnum_prop)&!is.na(temp)&
                                     !is.na(moist)&!is.na(nutrient)&
                                     !is.na(fire)&!is.na(dry))$age, plot = T)
testTemporalAutocorrelation(simulateResiduals(mod_abund_Erio_nozi),
                            subset(data_peat,
                                   Erio_prop>0&!is.na(temp)&
                                     !is.na(moist)&!is.na(nutrient)&
                                     !is.na(fire)&!is.na(dry))$age, plot = T)
testTemporalAutocorrelation(simulateResiduals(mod_abund_Erica),
                            subset(data_peat,
                                   !is.na(Erica_prop)&!is.na(temp)&
                                     !is.na(moist)&!is.na(nutrient)&
                                     !is.na(fire)&!is.na(dry))$age, plot = T)
testTemporalAutocorrelation(simulateResiduals(mod_abund_Carex),
                            subset(data_peat,
                                   !is.na(Carex_prop)&!is.na(temp)&
                                     !is.na(moist)&!is.na(nutrient)&
                                     !is.na(fire)&!is.na(dry))$age, plot = T)
```

Eriophprum and Ericaceae had significant temporal autocorrelation and will be refitted with an autoregressive AR covariance structure

#### Individual Sphagnum species

```{r}
testTemporalAutocorrelation(simulateResiduals(mod_abund_Medium),
                            subset(data_peat,
                                   !is.na(Medium/100)&!is.na(temp)&
                                     !is.na(moist)&!is.na(nutrient)&
                                     !is.na(fire)&!is.na(dry))$age, plot = T)
testTemporalAutocorrelation(simulateResiduals(mod_abund_Fuscum),
                            subset(data_peat,
                                   !is.na(Fuscum/100)&!is.na(temp)&
                                     !is.na(moist)&!is.na(nutrient)&
                                     !is.na(fire)&!is.na(dry))$age, plot = T)
testTemporalAutocorrelation(simulateResiduals(mod_abund_Rubellum),
                            subset(data_peat,
                                   !is.na(Rubellum/100)&!is.na(temp)&
                                     !is.na(moist)&!is.na(nutrient)&
                                     !is.na(fire)&!is.na(dry))$age, plot = T)
testTemporalAutocorrelation(simulateResiduals(mod_abund_Cuspidatum),
                            subset(data_peat,
                                   !is.na(Cuspidatum/100)&!is.na(temp)&
                                     !is.na(moist)&!is.na(nutrient)&
                                     !is.na(fire)&!is.na(dry))$age, plot = T)
```

### Autoregressive models

#### Erica

```{r}
data_peat_noNapredictors_Erica<-subset(data_peat,!is.na(temp)&!is.na(moist)&
                                   !is.na(nutrient)&!is.na(fire)&!is.na(dry))
times_Erica <- factor(data_peat_noNapredictors_Erica$age)
group_Erica <- factor(rep(1,101))

dat0_Erica <- data.frame(Erica_prop=data_peat_noNapredictors_Erica$Erica_prop,
                   temp=data_peat_noNapredictors_Erica$temp,
                   age=data_peat_noNapredictors_Erica$age,
                   moist=data_peat_noNapredictors_Erica$moist,
                   nutrient=data_peat_noNapredictors_Erica$nutrient,
                   fire=data_peat_noNapredictors_Erica$fire,
                   dry=data_peat_noNapredictors_Erica$dry,
                   fen=data_peat_noNapredictors_Erica$fen,
                   times_Erica, group_Erica)
```

```{r}
mod_abund_Erica_AR<-glmmTMB(Erica_prop~temp+moist+nutrient+
                                        fire+dry+ar1(times_Erica+0|group_Erica),
                                      family="beta_family",ziformula=~., 
                                      data=dat0_Erica)
summary(mod_abund_Erica_AR)
check_collinearity(mod_abund_Erica_AR)
```

#### Erio

```{r}
data_peat_noNapredictors_Erio<-subset(data_peat,!is.na(temp)&!is.na(moist)&
                                   !is.na(nutrient)&!is.na(fire)&!is.na(dry)&
                                     Erio_prop>0)
times_Erio <- factor(data_peat_noNapredictors_Erio$age)
group_Erio <- factor(rep(1,100))

dat0_Erio <- data.frame(Erio_prop=data_peat_noNapredictors_Erio$Erio_prop,
                        temp=data_peat_noNapredictors_Erio$temp,
                        age=data_peat_noNapredictors_Erio$age,
                        moist=data_peat_noNapredictors_Erio$moist,
                        nutrient=data_peat_noNapredictors_Erio$nutrient,
                        fire=data_peat_noNapredictors_Erio$fire,
                        dry=data_peat_noNapredictors_Erio$dry,
                        fen=data_peat_noNapredictors_Erio$fen,
                        times_Erio, group_Erio)
```

```{r}
mod_abund_Erio_nozi_AR<-glmmTMB(Erio_prop~temp+moist+nutrient+
                                        fire+dry+ar1(times_Erio+0|group_Erio),
                                      family="beta_family",ziformula=~0, 
                                      data=dat0_Erio)
summary(mod_abund_Erio_nozi_AR)
check_collinearity(mod_abund_Erio_nozi_AR)
```

### Model fit and predictive power

#### R2

```{r}
r2_zeroinflated(mod_abund_tot_Sphagnum)$R2_adjusted
r2_ferrari(mod_abund_Erio_nozi_AR,correct_bounds=T)
r2_zeroinflated(mod_abund_Erica_AR)$R2_adjusted
r2_zeroinflated(mod_abund_Carex)$R2_adjusted
r2_zeroinflated(mod_abund_Medium)$R2_adjusted
r2_zeroinflated(mod_abund_Fuscum)$R2_adjusted
r2_zeroinflated(mod_abund_Rubellum)$R2_adjusted
r2_zeroinflated(mod_abund_Cuspidatum)$R2_adjusted
```

#### RMSE

```{r}
data_peat_noNapredictors<-subset(data_peat,!is.na(temp)&!is.na(moist)&
                                   !is.na(nutrient)&!is.na(fire)&!is.na(dry))
dat0 <- data.frame(tot_Sphagnum_prop=data_peat_noNapredictors$tot_Sphagnum_prop,
                   Erio_prop=data_peat_noNapredictors$Erio_prop,
                   Erica_prop=data_peat_noNapredictors$Erica_prop,
                   Carex_prop=data_peat_noNapredictors$Carex_prop,
                   temp=data_peat_noNapredictors$temp,
                   age=data_peat_noNapredictors$age,
                   moist=data_peat_noNapredictors$moist,
                   nutrient=data_peat_noNapredictors$nutrient,
                   fire=data_peat_noNapredictors$fire,
                   dry=data_peat_noNapredictors$dry,
                   fen=data_peat_noNapredictors$fen)
```


```{r}
RMSE_tot_Sphagnum<-RMSE(predict(mod_abund_tot_Sphagnum,newdata=dat0,
                                type="response"),dat0$tot_Sphagnum_prop)
RMSE_Erio<-RMSE(predict(mod_abund_Erio_nozi_AR,newdata=dat0_Erio,
                        type="response"),dat0_Erio$Erio_prop)
RMSE_Erica<-RMSE(predict(mod_abund_Erica_AR,newdata=dat0,
                                type="response"),dat0$Erica_prop)
RMSE_Carex<-RMSE(predict(mod_abund_Carex,newdata=dat0,
                         type="response"),dat0$Carex_prop)
RMSE_Medium<-RMSE(predict(mod_abund_Medium,
                            newdata=subset(data_peat,!is.na(Medium)&
                                             !is.na(temp)&!is.na(moist)&
                                             !is.na(nutrient)&!is.na(fire)&
                                             !is.na(dry)),
                            type="response"),
                    subset(data_peat,!is.na(Medium)&!is.na(temp)&
                             !is.na(moist)&!is.na(nutrient)&!is.na(fire)&
                             !is.na(dry))$Medium)
RMSE_Fuscum<-RMSE(predict(mod_abund_Fuscum,
                            newdata=subset(data_peat,!is.na(Fuscum)&
                                             !is.na(temp)&!is.na(moist)&
                                             !is.na(nutrient)&!is.na(fire)&
                                             !is.na(dry)),
                            type="response"),
                    subset(data_peat,!is.na(Fuscum)&!is.na(temp)&
                             !is.na(moist)&!is.na(nutrient)&!is.na(fire)&
                             !is.na(dry))$Fuscum)
RMSE_Rubellum<-RMSE(predict(mod_abund_Rubellum,
                            newdata=subset(data_peat,!is.na(Rubellum)&
                                             !is.na(temp)&!is.na(moist)&
                                             !is.na(nutrient)&!is.na(fire)&
                                             !is.na(dry)),
                            type="response"),
                    subset(data_peat,!is.na(Rubellum)&!is.na(temp)&
                             !is.na(moist)&!is.na(nutrient)&!is.na(fire)&
                             !is.na(dry))$Rubellum)
RMSE_Cuspidatum<-RMSE(predict(mod_abund_Cuspidatum,
                            newdata=subset(data_peat,!is.na(Cuspidatum)&
                                             !is.na(temp)&!is.na(moist)&
                                             !is.na(nutrient)&!is.na(fire)&
                                             !is.na(dry)),
                            type="response"),
                    subset(data_peat,!is.na(Cuspidatum)&!is.na(temp)&
                             !is.na(moist)&!is.na(nutrient)&!is.na(fire)&
                             !is.na(dry))$Cuspidatum)
RMSE_tot_Sphagnum
RMSE_Erio
RMSE_Erica
RMSE_Carex
RMSE_Medium
RMSE_Fuscum
RMSE_Rubellum
RMSE_Cuspidatum
```

#### MAE

```{r}
MAE_tot_Sphagnum<-MAE(predict(mod_abund_tot_Sphagnum,newdata=dat0,
                                type="response"),dat0$tot_Sphagnum_prop)
MAE_Erio<-MAE(predict(mod_abund_Erio_nozi_AR,newdata=dat0_Erio,
                        type="response"),dat0_Erio$Erio_prop)
MAE_Erica<-MAE(predict(mod_abund_Erica_AR,newdata=dat0,
                        type="response"),dat0$Erica_prop)
MAE_Carex<-MAE(predict(mod_abund_Carex,newdata=dat0,
                         type="response"),dat0$Carex_prop)
MAE_Medium<-MAE(predict(mod_abund_Medium,
                            newdata=subset(data_peat,!is.na(Medium)&
                                             !is.na(temp)&!is.na(moist)&
                                             !is.na(nutrient)&!is.na(fire)&
                                             !is.na(dry)),
                            type="response"),
                    subset(data_peat,!is.na(Medium)&!is.na(temp)&
                             !is.na(moist)&!is.na(nutrient)&!is.na(fire)&
                             !is.na(dry))$Medium)
MAE_Fuscum<-MAE(predict(mod_abund_Fuscum,
                            newdata=subset(data_peat,!is.na(Fuscum)&
                                             !is.na(temp)&!is.na(moist)&
                                             !is.na(nutrient)&!is.na(fire)&
                                             !is.na(dry)),
                            type="response"),
                    subset(data_peat,!is.na(Fuscum)&!is.na(temp)&
                             !is.na(moist)&!is.na(nutrient)&!is.na(fire)&
                             !is.na(dry))$Fuscum)
MAE_Rubellum<-MAE(predict(mod_abund_Rubellum,
                            newdata=subset(data_peat,!is.na(Rubellum)&
                                             !is.na(temp)&!is.na(moist)&
                                             !is.na(nutrient)&!is.na(fire)&
                                             !is.na(dry)),
                            type="response"),
                    subset(data_peat,!is.na(Rubellum)&!is.na(temp)&
                             !is.na(moist)&!is.na(nutrient)&!is.na(fire)&
                             !is.na(dry))$Rubellum)
MAE_Cuspidatum<-MAE(predict(mod_abund_Cuspidatum,
                            newdata=subset(data_peat,!is.na(Cuspidatum)&
                                             !is.na(temp)&!is.na(moist)&
                                             !is.na(nutrient)&!is.na(fire)&
                                             !is.na(dry)),
                            type="response"),
                    subset(data_peat,!is.na(Cuspidatum)&!is.na(temp)&
                             !is.na(moist)&!is.na(nutrient)&!is.na(fire)&
                             !is.na(dry))$Cuspidatum)
MAE_tot_Sphagnum
MAE_Erio
MAE_Erica
MAE_Carex
MAE_Medium
MAE_Fuscum
MAE_Rubellum
MAE_Cuspidatum
```

## Models with interactions with fen factor

#### Carex

```{r}
mod_abund_Carex_nozi_int<-glmmTMB(Carex_prop~   
                               (temp+moist+
                               fire+dry)*fen+nutrient,  
                             family="beta_family", 
                             ziformula=~0,         
                             data=subset(data_peat,Carex/100>0)) 
summary(mod_abund_Carex_nozi_int)
```

### Model fit and predictive power

#### R2

```{r}
r2_ferrari(mod_abund_Carex_nozi_int,correct_bounds=T)
```

#### RMSE

```{r}
RMSE_Carex_ints<-RMSE(predict(mod_abund_Carex_nozi_int,
                              newdata=dat0,
                              type="response"),dat0$Carex_prop)
RMSE_Carex_ints
```

#### MAE

```{r}
MAE_Carex_ints<-MAE(predict(mod_abund_Carex_nozi_int,
                              newdata=dat0,
                              type="response"),dat0$Carex_prop)
MAE_Carex_ints
```

## Plots species groups

Plot Tot_Sphagnum Conditional Moisture:

```{r}
plot_tot_sphagnum_c_moist<-
  ggplot()+
  geom_point(data=data_peat,aes(x=moist,y=tot_Sphagnum_prop),
             size=3,
             alpha=0.2,
             shape=16)+
  geom_ribbon(data=
                data.frame(ggpredict(mod_abund_tot_Sphagnum,
                                     type="fixed",
                                     terms=c("moist[all]"))),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              color="grey",alpha=0.3)+
  geom_line(data=
              data.frame(ggpredict(mod_abund_tot_Sphagnum,
                                   type="fixed",
                                   terms=c("moist[all]"))),
            aes(x=x,y=predicted))+
  my_theme()+
  xlab("Increased local dry conditions")+
  ylab("Sphagnum abundance")
plot_tot_sphagnum_c_moist
```

Plot Tot_Sphagnum ZI Moisture:

```{r}
plot_tot_sphagnum_zi_moist<-ggplot()+
  geom_point(data=data_peat,aes(x=moist,y=pres_tot_Sphagnum),
             size=3,alpha=0.2,shape=16)+
  geom_ribbon(data=data.frame(ggpredict(mod_abund_tot_Sphagnum,
                                        type="zi_prob",terms=c("moist[all]"))),
              aes(x=x,y=1-predicted,ymin=1-conf.low,ymax=1-conf.high),
              color="grey",alpha=0.3)+
  geom_line(data=data.frame(ggpredict(mod_abund_tot_Sphagnum,
                                      type="zi_prob",terms=c("moist[all]"))),
            aes(x=x,y=1-predicted))+
  my_theme()+xlab("Increased local dry conditions")+ylab("Probability of presence Sphagnum")
plot_tot_sphagnum_zi_moist
```

Plot Erica ZI Moisture:

```{r}
plot_tot_erica_zi_moist <- ggplot()+
  geom_point(data=data_peat,aes(x=moist,y=pres_Erica),size=3,alpha=0.2,shape=16)+
  geom_ribbon(data=data.frame(ggpredict(mod_abund_Erica_AR,type="zi_prob",
                                        terms=c("moist[all]"))),
              aes(x=x,y=1-predicted,ymin=1-conf.low,ymax=1-conf.high),
              color="grey",alpha=0.3)+
  geom_line(data=data.frame(ggpredict(mod_abund_Erica_AR,type="zi_prob",
                                      terms=c("moist[all]"))),
            aes(x=x,y=1-predicted))+
  my_theme()+xlab("Increased local dry conditions")+ylab("Probability of presence Ericaceae")
plot_tot_erica_zi_moist
```

Plot Erio Conditional Temperature:

```{r}
plot_erio_c_temp<-ggplot()+
    geom_point(data=data_peat,aes(x=temp,y=Erio_prop),
               size=3,alpha=0.2,shape=16)+
    geom_ribbon(data=data.frame(ggpredict(mod_abund_Erio_nozi_AR,
                                          terms=c("temp[all]"))),
                aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
                color="grey",alpha=0.3)+
    geom_line(data=data.frame(ggpredict(mod_abund_Erio_nozi_AR,
                                        terms=c("temp[all]"))),
              aes(x=x,y=predicted))+
    my_theme()+xlab("Temperature")+ylab("Eriophorum abundance")
plot_erio_c_temp
```

Plot Erio Conditional Moisture:

```{r}
plot_erio_c_moist<-ggplot()+
  geom_point(data=data_peat,aes(x=moist,y=Erio_prop),
             size=3,alpha=0.2,shape=16)+
  geom_ribbon(data=data.frame(ggpredict(mod_abund_Erio_nozi_AR,
                                        terms=c("moist[all]"))),
                aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              color="grey",alpha=0.3)+
  geom_line(data=data.frame(ggpredict(mod_abund_Erio_nozi_AR,
                                      terms=c("moist[all]"))),
            aes(x=x,y=predicted))+
  my_theme()+xlab("Increased local dry conditions")+
  ylab("Eriophorum abundance") +
  theme(axis.title.y = element_text(color = "white"), 
        axis.text.y  = element_text(color = "white"))
plot_erio_c_moist
```

Plot Erio Conditional Fire:

```{r}
plot_erio_c_fire<-ggplot()+
    geom_jitter(data=data_peat,aes(x=fire,y=Erio_prop),
                position = position_jitter(0.1,0.01),
                size=3,alpha=0.2,shape=16)+
  geom_point(data=data.frame(ggemmeans(mod_abund_Erio_nozi_AR,
                                       type="fixed",terms=c("fire"))),
             aes(x=x,y=predicted),size=4,shape=16)+
  geom_errorbar(data=data.frame(ggemmeans(mod_abund_Erio_nozi_AR,
                                          type="fixed",terms=c("fire"))),
                aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
                width=0.2,size=0.5)+
    my_theme()+xlab("Fire activity")+ylab("Eriophorum abundance") +
  theme(axis.title.y = element_text(color = "white"), 
        axis.text.y  = element_text(color = "white"))
plot_erio_c_fire
```

Plot Carex Conditional Nutrient:

```{r}
plot_carex_c_nutrient<-ggplot()+
  geom_jitter(data=data_peat,aes(x=nutrient,y=Carex_prop),
              position = position_jitter(0.1,0.01),size=3,alpha=0.2,shape=16)+
  geom_point(data=data.frame(ggemmeans(mod_abund_Carex,
                                       type="fixed",terms=c("nutrient"))),
             aes(x=x,y=predicted),size=4,shape=16)+
  geom_errorbar(data=data.frame(ggemmeans(mod_abund_Carex,
                                          type="fixed",terms=c("nutrient"))),
                aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
                width=0.2,size=0.5)+
  my_theme()+xlab("External nutrient input")+ylab("Carex abundance")
plot_carex_c_nutrient
```

Plot Carex Conditional Fire:

```{r}
plot_carex_c_fire<-ggplot()+
  geom_jitter(data=data_peat,aes(x=fire,y=Carex_prop),
              position = position_jitter(0.1,0.01),size=3,alpha=0.2,shape=16)+
  geom_point(data=data.frame(ggemmeans(mod_abund_Carex,
                                       type="fixed",terms=c("fire"))),
             aes(x=x,y=predicted),size=4,shape=16)+
  geom_errorbar(data=data.frame(ggemmeans(mod_abund_Carex,
                                          type="fixed",terms=c("fire"))),
                aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
                width=0.2,size=0.5)+
  my_theme()+xlab("Fire activity")+ylab("Carex abundance") +
  theme(axis.title.y = element_text(color = "white"), 
        axis.text.y  = element_text(color = "white"))
plot_carex_c_fire
```

```{r}
plot_carex_c_int_fen_fire<-ggplot()+
  geom_jitter(data=data_peat,aes(x=fen,y=Carex_prop,
                                 color=fire),
              position=position_jitterdodge(jitter.width=0.1,
                                            jitter.height=0.01,
                                            dodge.width=0.5),
              size=3,alpha=0.2,shape=16)+
  geom_point(data=data.frame(ggemmeans(mod_abund_Carex_nozi_int,
                                        type="fixed",terms=c("fen","fire"))),
              aes(x=x,y=predicted,
                  color=group),
             size=4,shape=16,position=position_dodge(0.5))+
  geom_errorbar(data=data.frame(ggemmeans(mod_abund_Carex_nozi_int,
                                      type="fixed",terms=c("fen","fire"))),
            aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high,
                color=group),
            width=0.2,size=0.5,position=position_dodge(0.5))+
  scale_color_manual(values=c("#fdae61","#d7191c"))+
  my_theme_legend()+xlab("Fen")+ylab("Carex abundance")+
  labs(color="Fire activity") +
  theme(axis.title.y = element_text(color = "white"), 
        axis.text.y  = element_text(color = "white"),
        legend.position = c(0.25, 0.8))
plot_carex_c_int_fen_fire
```

### Figure 4

```{r}
fig4 <- (
  (plot_tot_sphagnum_c_moist) +
  plot_tot_sphagnum_zi_moist +
  (plot_tot_erica_zi_moist) +
  plot_erio_c_temp +
  plot_erio_c_moist +
  plot_erio_c_fire +
  (plot_carex_c_nutrient) +
  plot_carex_c_fire +
  plot_carex_c_int_fen_fire
) + plot_layout(ncol = 3, nrow = 3) +
  plot_annotation(tag_levels = "A", tag_prefix = "", tag_suffix = ")")
fig4
ggsave(filename="figures/fig4.tiff",plot=fig4,
       width=32,height=28,units="cm",dpi=300)
```

## Plots individual Sphagnum species

Plot Fuscum Conditional temp:

```{r}
plot_fuscum_c_temp<-ggplot()+
  geom_point(data=data_peat,aes(x=temp,y=Fuscum),size=3,alpha=0.2,shape=16)+
  geom_ribbon(data=data.frame(ggpredict(mod_abund_Fuscum,type="fixed",
                                        terms=c("temp[all]"))),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              color="grey",alpha=0.3)+
  geom_line(data=data.frame(ggpredict(mod_abund_Fuscum,type="fixed",
                                      terms=c("temp[all]"))),
            aes(x=x,y=predicted))+
  my_theme()+xlab("Temperature")+ylab("S. fuscum abundance")
plot_fuscum_c_temp
```

Plot Fuscum ZI temp:

```{r}
# Create variable for Fuscum presence/absence
data_peat$pres_Fuscum<-with(data_peat,ifelse(Fuscum>0,1,0))
plot_fuscum_zi_temp<-ggplot()+
  geom_point(data=data_peat,aes(x=temp,y=pres_Fuscum),size=3,alpha=0.2,shape=16)+
  geom_ribbon(data=data.frame(ggpredict(mod_abund_Fuscum,type="zi_prob",
                                        terms=c("temp[all]"))),
              aes(x=x,y=1-predicted,ymin=1-conf.low,ymax=1-conf.high),
              color="grey",alpha=0.3)+
  geom_line(data=data.frame(ggpredict(mod_abund_Fuscum,type="zi_prob",
                                      terms=c("temp[all]"))),
            aes(x=x,y=1-predicted))+
  my_theme()+xlab("Temperature")+ylab("Probability of S. fuscum presence")
plot_fuscum_zi_temp
```

Plot Fuscum ZI Fire:

```{r}
# Create average prediction (average line among factors=0 and factors=1)
average_prediction_fuscum_fire<-rbind(
  tibble(ggpredict(mod_abund_Fuscum,type="zi_prob",
                   terms=c("fire"),
                   condition=c(nutrient=1,dry=1)))%>%
    select(-group)%>%mutate(type="a"),
  tibble(ggpredict(mod_abund_Fuscum,type="zi_prob",
                   terms=c("fire"),
                   condition=c(nutrient=0,dry=0)))%>%
    select(-group)%>%mutate(type="b"),
  tibble(ggpredict(mod_abund_Fuscum,type="zi_prob",
                   terms=c("fire"),
                   condition=c(nutrient=0,dry=1)))%>%
    select(-group)%>%mutate(type="c"),
  tibble(ggpredict(mod_abund_Fuscum,type="zi_prob",
                   terms=c("fire"),
                   condition=c(nutrient=1,dry=0)))%>%
    select(-group)%>%mutate(type="d")
  )%>%
  group_by(x)%>%summarise(predicted=mean(predicted),
                          std.error=mean(std.error,na.rm=T),
                          conf.low=mean(conf.low,na.rm=T),
                          conf.high=mean(conf.high,na.rm=T))
plot_fuscum_zi_fire<-ggplot()+
  geom_jitter(data=data_peat,aes(x=fire,y=pres_Fuscum),
              position = position_jitter(0.1,0.01),size=3,alpha=0.2,shape=16)+
  # Use the average prediction as data
  geom_point(data=average_prediction_fuscum_fire,
             aes(x=x,y=1-predicted),size=4,shape=16)+
  geom_errorbar(data=average_prediction_fuscum_fire,
                aes(x=x,y=1-predicted,ymin=1-conf.low,ymax=1-conf.high),
                width=0.2,size=0.5)+
  my_theme()+xlab("Fire activity")+ylab("Probability of S. fuscum presence") +
  theme(axis.title.y = element_text(color = "white"), 
        axis.text.y  = element_text(color = "white"))
plot_fuscum_zi_fire
```

Plot Medium Conditional Nutrient:

```{r}
plot_medium_c_nutrient<-ggplot()+
  geom_jitter(data=data_peat,aes(x=nutrient,y=Medium),
              position = position_jitter(0.1,0.01),size=3,alpha=0.2,shape=16)+
  geom_point(data=data.frame(ggemmeans(mod_abund_Medium,
                                       type="fixed",terms=c("nutrient"))),
             aes(x=x,y=predicted),size=4,shape=16)+
  geom_errorbar(data=data.frame(ggemmeans(mod_abund_Medium,
                                          type="fixed",terms=c("nutrient"))),
                aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
                width=0.2,size=0.5)+
  my_theme()+xlab("External nutrient input")+ylab("S. medium abundance")
plot_medium_c_nutrient
```

Plot Medium ZI temp:

```{r}
# Create variable for Medium presence/absence
data_peat$pres_Medium<-with(data_peat,ifelse(Medium>0,1,0))
plot_medium_zi_temp<-ggplot()+
  geom_point(data=data_peat,aes(x=temp,y=pres_Medium),size=3,alpha=0.2,shape=16)+
  geom_ribbon(data=data.frame(ggpredict(mod_abund_Medium,type="zi_prob",
                                        terms=c("temp[all]"))),
              aes(x=x,y=1-predicted,ymin=1-conf.low,ymax=1-conf.high),
              color="grey",alpha=0.3)+
  geom_line(data=data.frame(ggpredict(mod_abund_Medium,type="zi_prob",
                                      terms=c("temp[all]"))),
            aes(x=x,y=1-predicted))+
  my_theme()+xlab("Temperature")+ylab("Probability of S. medium presence")
plot_medium_zi_temp
```

Plot Medium ZI moist:

```{r}
plot_medium_zi_moist<-ggplot()+
  geom_point(data=data_peat,aes(x=moist,y=pres_Medium),size=3,alpha=0.2,shape=16)+
  geom_ribbon(data=data.frame(ggpredict(mod_abund_Medium,type="zi_prob",
                                        terms=c("moist[all]"))),
              aes(x=x,y=1-predicted,ymin=1-conf.low,ymax=1-conf.high),
              color="grey",alpha=0.3)+
  geom_line(data=data.frame(ggpredict(mod_abund_Medium,type="zi_prob",
                                      terms=c("moist[all]"))),
            aes(x=x,y=1-predicted))+
  my_theme()+xlab("Increased local dry conditions")+ylab("Probability of S. medium presence") +
  theme(axis.title.y = element_text(color = "white"), 
        axis.text.y  = element_text(color = "white"))
plot_medium_zi_moist
```

Plot Cuspidatum Conditional Nutrient:

```{r}
plot_cuspidatum_c_nutrient<-ggplot()+
  geom_jitter(data=data_peat,aes(x=nutrient,y=Cuspidatum),
              position = position_jitter(0.1,0.01),size=3,alpha=0.2,shape=16)+
  geom_point(data=data.frame(ggemmeans(mod_abund_Cuspidatum,
                                       type="fixed",terms=c("nutrient"))),
             aes(x=x,y=predicted),size=4,shape=16)+
  geom_errorbar(data=data.frame(ggemmeans(mod_abund_Cuspidatum,
                                          type="fixed",terms=c("nutrient"))),
                aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
                width=0.2,size=0.5)+
  my_theme()+xlab("External nutrient input")+ylab("S. cuspidata abundance")
plot_cuspidatum_c_nutrient
```

Plot Cuspidatum ZI temp:

```{r}
# Create variable for Cuspidatum presence/absence
data_peat$pres_Cuspidatum<-with(data_peat,ifelse(Cuspidatum>0,1,0))
plot_cuspidatum_zi_temp<-ggplot()+
  geom_point(data=data_peat,aes(x=temp,y=pres_Cuspidatum),size=3,alpha=0.2,shape=16)+
  geom_ribbon(data=data.frame(ggpredict(mod_abund_Cuspidatum,type="zi_prob",
                                        terms=c("temp[all]"))),
              aes(x=x,y=1-predicted,ymin=1-conf.low,ymax=1-conf.high),
              color="grey",alpha=0.3)+
  geom_line(data=data.frame(ggpredict(mod_abund_Cuspidatum,type="zi_prob",
                                      terms=c("temp[all]"))),
            aes(x=x,y=1-predicted))+
  my_theme()+xlab("Temperature")+ylab("Probability of S. cuspidata presence")
plot_cuspidatum_zi_temp
```


### Figure 5

```{r}
fig5 <- (
  (plot_fuscum_c_temp) +
  plot_fuscum_zi_temp +
  (plot_fuscum_zi_fire) +
  plot_medium_c_nutrient +
  plot_medium_zi_temp +
  plot_medium_zi_moist +
  plot_cuspidatum_c_nutrient +
  plot_cuspidatum_zi_temp
) + plot_layout(ncol = 3, nrow = 3) +
  plot_annotation(tag_levels = "A", tag_prefix = "", tag_suffix = ")")
fig5
ggsave(filename="figures/fig5.tiff",plot=fig5,
       width=32,height=28,units="cm",dpi=300)
```

# Session info

```{r}
sessionInfo()
```

